### 2015-09-14 manuscript.Rnw
reworked intro, fix wrapping width

Phylogenetic information data is being used in many disciplines of the
life sciences, and phylogenetic comparative methods have seen their
use diversified in recent years (community phylogenetics,
...). However, phylogenetic information is scattered, often only
available as image files within publications \cite{Mounce}, and the
variety of file formats used to store phylogenetic information makes
it difficult for researchers to access

The Open Tree of Life (http://opentreeoflife.org) aims at assembling
and synthesizing the current understanding of phylogenetic
relationships across all organisms on Earth \cite{Hinchliff2015},
while providing tools and services to allow users to access this
information easily. OTL combines taxonomic information that serves as
the backbone for the phylogenetic relationships, and published
phylogenies to elucidate relationships among taxa.

The current draft of the Open Tree of Life contains 2.3 million
tips. Because of its comprehensive nature, researchers have access to
phylogenetic information for species that have yet to be included in
any phylogenetic tree as their position is inferred based on
taxonomy. Beyond obvious applications across the life sciences to
explore research questions in evolution, biodiversity, and
conservation, the resources OTL provides could be used in education
and outreach (e.g., illustrating course material, develop outreach
activities to explore relationships among species).

The Open Tree of Life project provides a web interface to explore,
query and download parts of the Open Tree
(http://tree.opentreeoflife.org). Additionally, the project provides
an Application Programming Interface (API) to the different services
it offers allowing users to interact programmatically with the
data. Here, we present \texttt{rotl} an R package that allows users to
interact with the OTL's API to search for and import phylogenetic
trees and taxonomic information. Having access to this data directly
from within R allows researchers to take advantage of the multiple R
packages that are available to visualize, analyze and manipulate
phylogenetic trees (ape, phytools, geiger, ggtree, phylobase,
etc... cite task view).

>>>>

Generating phylogenies has been facilitated by advances in sequencing and
computing technologies enabling the investigation of evolutionary relationships
among taxa. In parallel, new phylogenetic comparative methods have been
developed and use these phylogenies to explore fundamental questions about the
origin of biodiversity including the evolution of morphological and ecological
traits, the spatio-temporal variation in speciation rates, or both
\cite{Pennell2013}. The development of these methods has contributed to the need
for a greater availability of high-quality phylogenetic information across
life. However, phylogenetic information is scattered, often only available as
image files within publications, and the variety of file formats used to store
phylogenetic information makes it difficult for researchers to access and
integrate into their own research (\cite{Stoltzfus2012}, but see
\cite{Cranston2014} for suggestions of best practices).

The Open Tree of Life (OTL) project aims at assembling and synthesizing our
current understanding of phylogenetic relationships across all organisms on
Earth \cite{Hinchliff2015}, while providing tools and services that facilitate
access to this information. OTL combines taxonomic information that serves as
the backbone for the phylogenetic relationships, and published phylogenies to
elucidate relationships among taxa. This combination of information is used to
form the synthetic tree. Studies can be contributed to the synthetic tree
through the curator interface \url{https://tree.opentreeoflife.org/curator},
allowing the tree to be updated continuously. The current draft of the Open Tree
of Life contains 2.3 million tips. Beyond obvious applications across the life
sciences to explore research questions in evolution, biodiversity, and
conservation, the resources OTL provides could be used in education and outreach
(e.g., illustrating course material, develop outreach activities to explore
relationships among species).

% Because of its comprehensive nature, researchers have access to phylogenetic
% information for species that have yet to be included in any phylogenetic tree
% as their position is inferred based on taxonomy.

The Open Tree of Life project provides a web interface to explore, query and
download parts of the synthetic Open Tree
(\url{http://tree.opentreeoflife.org}). Additionally, the project provides an
Application Programming Interface (API) to the different services it offers
allowing users to work programmatically with the underlying data. Here, we
present \texttt{rotl} an R package that allows users to interact with the OTL's
API to search for and import phylogenetic trees and taxonomic
information. Having access to this data directly from within R allows
researchers to take advantage of the multiple R packages that are available to
visualize, analyze and manipulate phylogenetic trees (ape, phytools, geiger,
ggtree, phylobase, etc... cite task view).
